---
title: 'Assignment 2: Data visualization'
author: "Achilles Isidore Israel"
output: html_document
editor_options: 
  chunk_output_type: console
---

You will have to create 3 plots based on the datasets and instructions detailed below. You will find the plots themeselves in the `assignments/assignment_2_plots`. Your task is to write the code that will reproduce the plots as closely as possible.

# Skills needed to solve this assignment

-   Using R and RStudio, reading data
-   Reporting using RMarkdown
-   Using Git and Github (for submitting the task)
-   Data manipulation (e.g. dplyr, tidyr), and working with factors (forcats)
-   Data visuzlization (ggplot2)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(tidytuesdayR)
```

## Task 1: Climbing expeditions

The 2020-09-22 TidyTueday datasets are about climbing expeditions. From the three datasets, use the "expeditions". Reproduce the plot below! Notice a few things:

-   Use `forcats::fct_lump()` to get the 15 most frequent peaks, and drop the "Other" category.
-   The bars are ordered by the sum of all expeditions (use `fct_reorder()`).
-   The bar colors use the viridis palette and light theme.

```{r}

library(tidyverse)
library(tidytuesdayR)

tt_data <- tidytuesdayR::tt_load("2020-09-22")
expeditions <- tt_data$expeditions

##Prepare the data: top 15 peaks only, reorder by count
 expedition_counts <- expeditions %>%
  mutate(peak_name = fct_lump(peak_name, n = 15)) %>% 
  filter(peak_name != "Other") %>%            
  count(peak_name) %>%                        
  mutate(peak_name = fct_reorder(peak_name, n))                                                                     
##Plot
ggplot(expedition_counts, aes(x = peak_name, y = n, fill = peak_name)) +
  geom_col(show.legend = FALSE) +
  scale_fill_viridis_d() +
  theme_light() +
  labs(
    x = "Peak",
    y = "Number of Expeditions",
    title = "Top 15 Peaks by Number of Expeditions"
  ) +
  coord_flip()


```

## Task 2: PhDs awarded

The 2019-02-19 TidyTueday dataset is about phd-s awarded by year and field. There is only one dataset, it is called `phd_by_field`. Reproduce the plot below!

Notes:

-   First you have to aggregate the data to count all phd-s by broad fields.
-   To make the x axis breaks pretty, use `scales::pretty_breaks()`, to make the y axis labels comma formatted, use `scales::comma_format()`.
-   The line size is 1.2, the colors are from the brewer "Dark2" palette. The theme is set to minimal.

```{r}

library(tidyverse)
library(tidytuesdayR)
library(scales)

phd <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-02-19/phd_by_field.csv")

##Summarize total PhDs by year and broad field
phd_summary <- phd %>%
  group_by(year, broad_field) %>%
  summarise(total_phds = sum(n_phds, na.rm = TRUE), .groups = "drop")

##Plot
ggplot(phd_summary, aes(x = year, y = total_phds, color = broad_field)) +
  geom_line(size = 1.2) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(labels = comma_format()) +
  scale_color_brewer(palette = "Dark2") +
  theme_minimal() +
  labs(
    title = "PhDs Awarded by Broad Field Over Time",
    x = "Year",
    y = "Number of PhDs",
    color = "Field"
  )

```

## Task 3: Commute in the US

The 2019-11-05 TidyTueday dataset is about commuting to work in each city in the US by bike or on foot. There is only one dataset, it is called `commute`. Reproduce the plot below!

Notes:

-   First you have to aggregate the data to count all commutes by state.
-   Both axis scales are log transformed and the labels comma formatted, using `scales::comma_format()`
-   The point size is 2, . The theme is set to light.
  
```{r}

library(tidyverse)
library(scales)
library(tidytuesdayR)
library(dplyr)
library(tidyr)

commute <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-11-05/commute.csv")

##Aggregate the data by state
commute_by_state_mode <- commute %>%
  filter(mode %in% c("Bike", "Walk")) %>%
  group_by(state, mode) %>%
  summarise(total = sum(n, na.rm = TRUE), .groups = "drop")

commute_by_state_wide <- commute_by_state_mode %>%
  pivot_wider(names_from = mode, values_from = total)

##Plot

commute_by_state_wide_filtered <- commute_by_state_wide %>%
  filter(!is.na(Bike), !is.na(Walk), Bike > 0, Walk > 0)

ggplot(commute_by_state_wide_filtered, aes(x = Bike, y = Walk)) +
  geom_point(size = 2) +
  scale_x_log10(labels = comma_format()) +
  scale_y_log10(labels = comma_format()) +
  labs(
    title = "Bike vs Walk Commuters by State",
    x = "Total Bike Commuters (log scale)",
    y = "Total Walk Commuters (log scale)"
  ) +
  theme_light()



```
