---
title: "FinalProject"
author: "Achilles Isidore Israel"
date: "2025-05-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read the dataset of Pixar Films

```{r}

pixar_films <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-11/pixar_films.csv')
public_response <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-11/public_response.csv')
box_office <- readr::read_csv('https://raw.githubusercontent.com/erictleung/pixarfilms/master/data-raw/box_office.csv')
```

##Run an exploratory data analysis (EDA) to investigate the dataset.

```{r}

##Load necessary packages
library(tidyverse)
library(tidytuesdayR)

##View structure of datasets
view(pixar_films)
view(public_response)
view(box_office)


```

Fixing Coding Error

```{r}

##Remove column 27 in the Pixar_films dataset
pixar_films <- pixar_films[-27, ]

##Replace missing values in running time with 0 (not the mean)
pixar_films$run_time[is.na(pixar_films$run_time)] <- 0

##Replace missing values in Rotten Tomatoes with 0 (not the mean)
public_response$rotten_tomatoes[is.na(public_response$rotten_tomatoes)] <- 0

##Replace missing values in Metacritic with 0 (not the mean)
public_response$metacritic[is.na(public_response$metacritic)] <- 0

##Replace missing values in critics_choice with 0 (not the mean)
public_response$critics_choice[is.na(public_response$critics_choice)] <- 0

##Replace missing values in cinema_score with N/A
public_response$cinema_score[is.na(public_response$cinema_score)] <- "N/A"

##Ensure dates and numeric values are correctly typed
pixar_films$release_date <- as.Date(pixar_films$release_date)

pixar_films$run_time <- as.integer(pixar_films$run_time)

public_response$rotten_tomatoes <- as.integer(public_response$rotten_tomatoes)

public_response$metacritic <- as.integer(public_response$metacritic)

public_response$critics_choice <- as.integer(public_response$critics_choice)

##Match Film Names across datasets (Trim and Clean)
pixar_films$film <- trimws(pixar_films$film)
public_response$film <- trimws(public_response$film)

pixar_films$film <- tolower(pixar_films$film)
public_response$film <- tolower(public_response$film)

box_office_trimmed <- box_office %>%
  select(film, box_office_worldwide)

box_office_trimmed$film <- tolower(box_office_trimmed$film)
```

Make one data set adding pixar_films, public_responses, and box_office_worldwide of the box_office data set

```{r}

library(dplyr)

##Ensure Film columns are clean and aligned
pixar_films <- pixar_films %>%
  mutate(film = str_trim(film))

public_response <- public_response %>%
  mutate(film = str_trim(film))

##Join the three datasets
combined_data <- pixar_films %>%
  right_join(public_response, by = "film") %>%
  left_join(box_office_trimmed, by = "film")

```

Add column to find the average of rotten tomatoes, metacritic, and critic's choice 
```{r}

##Install and Load package/s
library(purrr)
library(dplyr)

##Find the average score
combined_data <- combined_data %>%
  mutate(
    average_public_response_score = pmap_dbl(
      list(rotten_tomatoes, metacritic, critics_choice),
      function(rt, mc, cc) {
        scores <- c(rt, mc, cc)
        valid_scores <- scores[scores > 0 & !is.na(scores)]
        if (length(valid_scores) == 0) NA else mean(valid_scores)}))

##Top 10
top_10 <- combined_data %>%
  arrange(desc(average_public_response_score)) %>%
  slice_head(n = 10)

```

Use ggplot for visualization for the Film and Public Response Scores (RT, MC, CC), and the Average Public Response Score but only showing the Top 10 

```{r}

##Load Packages
library(ggplot2)
library(tidyr)
library(dplyr)

##Get the top 10 films by average_score
top_10_films <- combined_data %>%
  arrange(desc(average_public_response_score)) %>%
  slice_head(n = 10) %>%
  pull(film)

##Filter combined_data to only top 10 films and arrange by descending average_score
filtered_data <- combined_data %>%
  filter(film %in% top_10_films) %>%
  arrange(desc(average_public_response_score)) %>%
  mutate(film = factor(film, levels = unique(film)))

##Split into two groups
films_first_5 <- levels(filtered_data$film)[1:5]
films_second_5 <- levels(filtered_data$film)[6:10]

##Prepare data for first 5 films
first_graph <- filtered_data %>%
  filter(film %in% films_first_5) %>%
  select(film, rotten_tomatoes, metacritic, critics_choice, average_public_response_score) %>%
  pivot_longer(
    cols = c(rotten_tomatoes, metacritic, critics_choice, average_public_response_score),
    names_to = "score_type",
    values_to = "score"
  ) %>%
  mutate(film = factor(film, levels = films_first_5))

##Prepare data for second 5 films
second_graph <- filtered_data %>%
  filter(film %in% films_second_5) %>%
  select(film, rotten_tomatoes, metacritic, critics_choice, average_public_response_score) %>%
  pivot_longer(
    cols = c(rotten_tomatoes, metacritic, critics_choice, average_public_response_score),
    names_to = "score_type",
    values_to = "score"
  ) %>%
  mutate(film = factor(film, levels = films_second_5))

##Plot the first 5 films
p1 <- ggplot(first_graph, aes(x = score, y = film, fill = score_type)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(
    title = "Top 5 Pixar Films by Public Rating (1-5)",
    x = "Score",
    y = "Film",
    fill = "Score Type"
  ) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10))

##Plot the next 5 films
p2 <- ggplot(second_graph, aes(x = score, y = film, fill = score_type)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(
    title = "Top 5 Pixar Films by Public Rating (6-10)",
    x = "Score",
    y = "Film",
    fill = "Score Type"
  ) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10))

print(p1)
print(p2)
```

Find the relationship between RT, MC, CC and box office performance.

Null Hypothesis: There is no relationship between RT, MC, CC, scores and worldwide box office revenue.
Alternative Hypothesis: Higher RT, MC, CC, scores are associated with higher worldwide box office revenue.

```{r}

##
library(zoo)

##Find the relationship
model <- lm(box_office_worldwide ~ rotten_tomatoes + metacritic + critics_choice, data = combined_data)
summary(model)

#Check Residual Diagnostic
par(mfrow = c(2, 2))
plot(model)

##Check Multicollinearity (Variance Inflation Factor)
library(car)
vif(model)

##Check Shapiro-Wilk Test for Normality of Residuals
shapiro.test(residuals(model))

##Check Breusch-Pagan Test for Heteroscedasticity
library(lmtest)
bptest(model)

##Check for Influential Observations (Cook's Distance)
plot(model, which = 4)


```
Test another hypothesis to check if there a difference in box office worldwide revenue between PG and G films?"

Null Hypothesis: There is no difference in mean worldwide box office revenue between PG and G films.
Alternative Hypothesis: There is a difference in mean worldwide box office revenue between PG and G films.

```{r}

library(dplyr)

pg_g_data <- combined_data %>%
  filter(film_rating %in% c("G", "PG"), !is.na(box_office_worldwide))

t.test(box_office_worldwide ~ film_rating, data = pg_g_data)

```
